### 2023-04-20 paper/paper.Rmd
first commit

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'rObsPack: An R package for the NOAA global observations ObsPack methane GLOBALVIEW+'
tags:
  - R
  - NOAA
  - CH4
  - methane
  - ObsPack
authors:
  - name: Sergio Ibarra-Espinosa
    orcid: 0000-0002-3162-1905
    equal-contrib: true
    affiliation: "1, 2"
  - name: Lei Hu
    orcid: 0000-0002-5624-3832
    equal-contrib: true
    affiliation: "2"
affiliations:
 - name: Cooperative Institute for Research in Environmental Sciences, University of Colorado Boulder, Boulder, CO, USA
   index: 1
 - name: NOAA Global Monitoring Laboratory, Boulder, CO, USA
   index: 2
citation_author: Ibarra-Espinosa and Hu
date: 04 April 2023
year: 2023
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---
<!-- output: rticles::joss_article -->
<!-- output: word_document -->

# Summary

Observations of greenhouse gas emissions are critical to monitor 
the state of the atmosphere, quantify present and historical 
emissions, and understand global climate change. Observation 
Package (ObsPack) is a specific format to deliver atmospheric
measurements data of trace gasses [@masarie2014obspack]. 
The NOAA Global Monitoring Laboratory (GML) manages a 
global network of atmospheric observations compiled and 
delivered to the public as ObsPack CH4 GLOBALVIEW+. This 
product provides access to global methane observations made 
in surface, tower, aircrafts, aircores and ships. Processing 
global data can be cumbersome since they are provided as 
multiple text and NetCDF files. The text files consist of a 
commented metadata followed with the tabulated data which can 
have different headers across files. The metadata can have critical
information which determines the nature of the file itself, such as
altitude of the site, local hour, laboratory and more. Below the
metadata, starts the section with tabulated data, with different
header across files.

# Statement of need

In order to analyze greenhouse gasses measurements we developed the 
R package `robspack`, which reads and processes NOAA/GML ObsPack
GLOBALVIEW+ CH4 data. This package imports functions from 
`data.table` R package, which contains C bindings with 
parallel implementation via Open-MP [@dt]. 
`data.table` is faster than other Python, Julia and R 
implementations for data-science, providing a strong basis
for `robspack`. The main objective of `robspack` is to
integrate different CH4 ObsPack text files in a long 
format, adding fields for specific metadata information,
with a tidy structure [@Silge2016]. Hence, the data 
can easily be processed with `data-table`, allowing
to produce plots with `ggplot2` and other R packages [@g2].

The steps followed by `robspack` are described in the 
following lines. First, it is necessary to create a 
summary of the data , read all the file names, identify 
key-words, sectors and full paths. The data is read
iterating on each file for each category such as aircraft, tower, 
surface, flasks, ships and aircore. It is necessary to
add UTC, local time and calculate altitude above ground 
level (agl) from the data. For continuous ground sites, 
afternoon observations are selected and averaged to reduce
correlations and noise. Given the high frequency of 
observations of data, usually every second, aircraft data is
usually averaged every 20 seconds. Once the data is processed
for each category, it is merged in a long format. The 
installation and functions created to perform these tasks 
are shown in the following sections.

# Functions

The `robspack` functions are shown in the table 1.

Table 1. Functions and classes in `robspack`.

| Function              | Description                                       |
|-----------------------|---------------------------------------------------|
| `invfile`             | Class with `print`, `summary` and `plot` methods  |
| `obs_addltime()`      | Add local time based on metadata and longitude    |
| `obs_addtime()`       | Add UTC time                                      |
| `obs_addzero()`       | Format numbers to match specific requirements     |
| `obs_agg()`           | Aggregates ObsPack by time                        |
| `obs_find_receptors()`| Find expected receptors and NetCDF files          |
| `obs_format()`        | Format for some columns of data.table             |
| `obs_freq()`          | Return numeric vector in intervals                |
| `obs_invfiles()`      | Construct `invfile` objects                       |
| `obs_list.dt()`       | Rbind list of data.frames with different names    |
| `obs_meta()`          | Reads ObsPack metadata                            |
| `obs_out()`           | Outersect, opposed as intersect                   |
| `obs_rbind()`         | Rbind data.frames with different names            |
| `obs_read()`          | Read files, and add metadata as columns           |
| `obs_read_csvy()`     | Read csvy file and prints yaml header             |
| `obs_roundtime()`     | Round seconds from "POSIXct" "POSIXt" classes     |
| `obs_summary()`       | Construct summary of ObsPack as a data.frame      |
| `obs_table()`         | Return a data.frame with summary of data          |
| `obs_trunc()`         | Trunc numbers with a desired number of decimals   |
| `obs_write()`         | Write CSVY to disk, YAML followed by tabulated    |

# Installation

To install `robspack`, the user must have installed the
R package `remotes` and run the following script. 
This process will install all the required 
dependencies, such as `data.table`, `cptcity`, an R package with
more than 7000 color palettes, and `lubridate`, a package to manage 
time and dates [@cpt;@lu].
Then, we call the libraries to load the
function into the environment.

```{r, eval = F}
remotes::install_github("ibarraespinosa/robspack")
library(robspack)
library(data.table)
```

```{r, include = F, message=F, warning=F}
library(robspack)
library(data.table)
```

For this manuscript, we are presenting
the application of `robspack` on tower observations.

## ObsPack summary

The first step consists in constructing a summary for the ObsPack.
This is required to read the data, but also, identify `agl`, 
which is present in some of the file names.
This function returns a `data.frame`.
Optionally, the user can indicate a path to store the `data.frame`.
`obs_summary` also prints a summary of the data. The second
argument is the categories, and by default includes the categories
shown below, to account for all the files.
Then the summary `data.frame` contains 
the columns `id` as the full path to each
file, `name` which is the name or relative path of the file, `n`
just an id, `sector` such as tower, and the column `agl`
which indicates the `agl` indicated in the name of the file
if available. To read the documentation of this function,
the user must run `?obs_summary`.

```{r}

categories <- c(
  "aircraft-pfp",
  "aircraft-insitu",
  "surface-insitu",
  "aircore",
  "surface-pfp",
  "tower-insitu",
  "shipboard-insitu",
  "flask"
)
obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
index <- obs_summary(obs = obs)

```

There are 362 files in the ObsPack directory.
The printed information also shows the total at the bottom,
as the sum of the individual file by sector. This
is to ensure that the sum of files is equal to the
total number of files found, shown at the top.
furthermore, the printed information
also shows that there are 136 files with the
`agl` explicitly mentioned in the name of the file.

Sometimes we need more information about the site. For instance,
what do the observations start and end. Then, we added 
the function `obs_table`, which calculates statistics
summary of "time" and other numeric variables by file name,
sector, site, altitude and mode. For instance,
the observations in the site "SCT" in
South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
"2020-12-31 23:30:00 UTC".
```{r}
dft <- obs_table(index = index,
                 categories = "tower-insitu",
                 verbose = FALSE)
dft[site_code == "SCT", ]$timeUTC |> 
  range()
```

### Read data

Once the summary is built, the function `obs_read`
will read the files available in the
index file previously generated.
Here we selected the category "tower-insitu".
The argument verbose prints which files are being read each time,
by default.
At the end, this function prints the total number
of observations by type of altitude (agl or asl).

```{r}
df <- obs_read(index = index,
               categories = "tower-insitu",
               verbose = FALSE)

```

We added a function to plot the data read from ObsPack.
The y-axis is the field `value` and the 
x-axis is by default `time`. The data illustrated sorted by
color is the field `site_code`, with the default
number of 3 sites. The argument `pal` is to define the
color palette, used by the internally imported function
`cptcity::cpt`.

```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
obs_plot(dt = df, time = "time", yfactor = 1e9, cex = 0.5)
```

Here we can see `r round(nrow(df)/1000000, 2)` million observations 
for `tower-insitu`. These observations are made between 
`r range(df$year)[1]` and `r range(df$year)[2]`. The identification
of the altitude and type is critical. The approach used here 
consists of:

1. Identify `agl` from the name of the tile.
2. If `agl` not present, search fill_values used in elevation and transform them into NA (not available)
3. If `agl` is not present, `agl = altitude - elevation`.
4. If there are some NA in elevation, will result some NA in `agl`
5. A new column is added named `altitude_final` to store `agl` or `asl`
6. Another column named `type_altitude` is added to identify `agl` or `asl`.
7. If there is any case NA in `altitude_final`, `type_altitude` is "not available"

### Filtering

ObsPack includes global observations and sometimes we need to extract data
for a specific region and periods of time.
In this part we include spatial and temporal parameters
to filter data.
The year of interest is 2020, but we also included
December of 2019 and January of 2021.
At this stage, we can apply the spatial filter
by using the coordinates.

```{r}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
evening <- 14:15

yy <- 2020
df <- rbind(df[year == yy - 1 & month == 12],
            df[year == yy],
            df[year == yy + 1 & month == 1])

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
```

After filtering by space and time, we have `r round(nrow(df), 2)`
million observations. Towers can have observations at different
heights. Here we need to select one site with the observations
registered at the highest height. The column
with the height is named `altitude_final` and the
max altitude was named `max_altitude`.

```{r}
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
```

### Time

Here we need to start time columns.
The function `obs_addtime` adds time columns
`timeUTC`, `timeUTC_start` which shows the start time
of each observation and `timeUTC_end`
which shows the end time for each observation.
Then we need to identify the local time with the function
`add_ltime`. This is important because to identifying
observations in the evening in local time.
`add_ltime` uses two methods, first identify
the time difference with utc by identifying the
metadata column "site_utc2lst". If this information
is not available, with the aircrafts for instance,
the local time is calculated with an approximation
based on longitude:

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where $lt$ is the local time, $UTC$ the time,
$longitude$ the coordinate.
Then, the time is cut every two hours.
Now, we identify the local time to select evening hours.

```{r}
df2 <- obs_addtime(df)
df2$timeUTC <- cut(x = df2$timeUTC+3600,
                   breaks = "2 hour") |>
  as.character() |>
  as.POSIXct(tz = "UTC")
df3 <- obs_addltime(df2)
df3 <- df3[lh %in% evening]
```

Now there are `r nrow(df3)` observations.
At this point we can calculate the averages of several columns by
the cut time. The function `obs_agg` does this aggregation as shown
in the following lines of code. The argument `gby`
establish the function used to aggregate `cols`, in this case
the function mean by time and altitude. Finally, we
add local time again.

```{r}
df4 <- obs_agg(dt = df3,
               gby = "mean",
               cols = c("value", "latitude", "longitude", "type_altitude",
                        "dif_time", "year_end", "site_utc2lst"),
               verbose = FALSE,
               byalt = TRUE)
df5 <- obs_addltime(df4)
```

Now there are `r nrow(df5)` observations,
`r nrow(df3) - nrow(df5)` less observations.
Here we add the column `max_altitude` to identify the max altitude
by site.

```{r}
df5[,
    max_altitude := max(altitude_final),
    by = site_code]
df5[,
    c("site_code",
      "altitude_final",
      "max_altitude")] |> unique()
```

### Saving master as text and csvy

Now that we have all the required information, 
we can save the files. Here, we name the data.frame 
as master, because it contains all the information.
This is important because some 
fields can be used in the future, and for traceability.
For convenience,
time variables are transformed into character
before writing into the disk.
The separation is space " ".

```{r, eval = T}
master <- df5
master$timeUTC <- as.character(master$timeUTC)
master$timeUTC_end <- as.character(master$timeUTC_end)
master$local_time <- as.character(master$local_time)

fwrite(master, 
       file = "tower_insitu_2020.txt",
       sep = " ")
```

The format Comma Separated Value with YAML (CSVY)[^3]
consists in a typical CSV with a YAML header.
The function`obs_write` includes the argument
`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write` adds 
the output of the R function `str`, which
provides a vertical summary of the data, known as
structure.

```{r, eval = T}
obs_write_csvy(dt = master,
              notes = "tower 2020",
              out = "tower_insitu_2020.csvy")
```

To check the YAML header we read the first 38 lines of 
the files that were generated. Here we can see
the column names, type of data and first observations.
The YAML header is delimited by the characters "---".

```{r, eval = F}
readLines("tower_insitu_2020.csvy")[1:38]
```

### Saving receptors

We need to filter some columns
from the master files in a new object called receptors.
This is needed because internally we run HYSPLIT [@hy]
using the information from the receptors.
In the case of a tower, we need to select 
observations with the highest 
altitude. The specific columns are selected
as shown on the following code. We are selecting the 
ending times, because later HYSPLIT is run backwards
based on the time of measurement, between ending and
starting times. The columns about time are formatted to
have two characters. For instance, the month 1,
is formatted as "01". We also need to filter for
`type_altitude` equal 0, representing `agl `observations
, or equal to 1, `asl`.

```{r, eval = F}

receptor <- master[altitude_final == max_altitude,
                   c("site_code",
                     "year", "month", "day",
                     "hour", "minute", "second",
                     "latitude", "longitude",
                     "altitude_final", "type_altitude",
                     "year_end", "month_end", "day_end", "hour_end",
                     "minute_end", "second_end")]
receptor$altitude_final <- round(receptor$altitude_final)
receptor <- obs_format(receptor)

if(nrow(receptor_agl) > 0) {
  fwrite(x = receptor_agl,
         file = "paper/receptor_tower_insitu_2020_AGL.txt"),
  sep = " ")}

if(nrow(receptor_asl) > 0) {
  fwrite(x = receptor_asl,
         file = "paper/receptor_tower_insitu_2020_ASL.txt"),
  sep = " ")}

```
## Recommendation for other applications
The approach to generate receptors depends on each type of 
observation and other considerations. For instance, aircraft with 
continuous observations at each second can be filtered and averaged 
every 20 seconds. In that way, the footprints are still 
representative and it would not be necessary to run HYSPLIT 
every second. Of course, it depends on the application 
and objective of the study. For this manuscript, we are 
presenting the generation of receptors based on tower observations. 
Furthermore, in this package we are sharing scripts to process other sectors
The scripts are available in the path 
`https://github.com/ibarraespinosa/robspack/tree/main/rscripts`

## Conclusion

In this manuscript we presented an `robskpack`,
an R package to read and process
$CH_4$ ObsPack GLOBALVIEW+ published by the Global Monitoring 
Laboratory (GML) from the National Oceanographic and Atmospheric
Administration (NOAA). `robspack` reads the text data
which have different headers and organizes them in a common 
format. Then, this software applies calculations to
filter observations by time and space. Finally,
this software generates receptors in a suitable format that 
allows it to run HYSPLIT and generate footprints. 
This software does 
not provide methods to run HYSPLIT, but the user
can follow the site https://www.ready.noaa.gov/HYSPLIT.php.

[^1]: https://h2oai.github.io/db-benchmark/
[^2]: https://gml.noaa.gov/ccgg/carbontracker-lagrange/
[^3]: https://csvy.org/

# Acknowledgements

Funding: This project is funded by the NOAA Climate Program Office AC4
and COM programs (NA21OAR4310233 / NA21OAR4310234). This research was 
supported by the NOAA cooperative agreement NA22OAR4320151.

```{r, include=FALSE}
file.copy(from = "paper.md", to = "../", overwrite = TRUE)
file.copy(from = "paper.bib", to = "../", overwrite = TRUE)
```

# References