### 2023-06-12 rscripts/00_README.Rmd
paper

---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### files:
```{r, echo=F}
list.files(recursive = T)
```

The directory `rscripts` contains the required files
to generate the list of receptors using `robspack`
The scripts will run from a defined working directory

### 1. Run index (summary)

The script `r/index.R` will create the tree directory
to store the processed ObsPack files
Also, it will create summaries and store them
for each category in the directory `obs`

### 2. Categories in the directory r
These scripts need to be run using Rscript command
examples:

- `Rscript r/aircore_year.R 2020 obs/obs_aircore.csv.gz`
- `Rscript r/aircraft_year.R 2020 obs/obs_aircraft-insitu.csv.gz`
- `Rscript r/flask_year.R 2020 obs/obs_flask.csv.gz`
- `Rscript r/surface_year.R 2020 obs/obs_surface-insitu.csv.gz`
- `Rscript r/tower_year.R 2020 obs/obs_tower-insitu.csv.gz`

Windows users can adapt the following lines

- "C:/Users/*User*/My Documents/R/R-4.1.3/bin/Rscript.exe" r/tower_insitu_year.R 2020 obs/obs_tower-insitu.csv.gz
- "C:/Users/*User*/My Documents/R/R-4.1.3/bin/Rscript.exe" r/aircore_year.R 2020 obs/obs_aircore.csv.gz

### 3. Run Hysplit
Once if we have the receptor, we can
run Hysplit to generate footprints in NetCDF format

### 4. Generate receptor info
Once we have the receptor list and the footprints were generated,
we can generate the receptor info, which
consists in text files.
The script is `r/inputs_inv.R`