### 2023-07-12 paper_elsevier/paper_elsevier.Rmd
add paper

---
title: robspack An R package for processing atmospheric greenhouse gas observations in NOAA's ObsPack
author:
  - name: Sergio Ibarra-Espinosa
    email: sergio.ibarra-espinosa@noaa.gov
    affiliation: a,b
    correspondingauthor: true
    footnote: 1
  - name: Lei Hu
    email: lhu@noaa.gov
    affiliation: b
address:
  - code: a
    organization: Cooperative Institute for Research in Environmental Sciences, CU Boulder
    addressline: 216 UCB, University of Colorado Boulder campus
    city: Boulder
    state: Colorado
    postcode: 80309
    country: United States
  - code: b
    organization: NOAA Global Monitoring Laboratory
    addressline: 325 Broadway
    state: Colorado
    postcode: 80309
    country: United States
footnote:
  - code: 1
    text: "This is the first author footnote."
abstract: |
  In this study, we present a new open-source R package `robspack`, 
  to read, process, select, and plot NOAA Observation Package (ObsPack)
  data products. We use a methane ObsPack data product as an example
  in this code base, but it can be easily modified to analyze ObsPack
  products for other greenhouse gasses. The R package starts with 
  creating a catalog of  all ObsPack files in each product. It then 
  reads all files and creates one database. While reading each ObsPack 
  file, it extracts site elevation and time zone information from the 
  file header and calculates sampling altitude in meters above ground 
  level and local time for individual samping events. Finally, it 
  processes and selects observations for inverse modeling purposes. 
  This package imports functions from data.table R package, which 
  contains C bindings with parallel implementation via Open-MP 
  [@dt]. data.table is faster than other 
  Python, Julia and R implementations for data-science, 
  providing a strong basis for robspack. robspack provides functions 
  to perform these tasks in a transparent and efficient way, 
  supporting open-source communities in environmental sciences.
keywords: 
  - ObsPack
  - NOAA
  - Greenhouse gases
journal: "Environmental Modelling & Software"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

<!-- output:  -->
<!--   rticles::elsevier_article: -->
<!--     keep_tex: true -->
<!--     citation_package: natbib -->

<!-- output: word_document -->
# Introduction

<!-- New Introduction -->
<!-- 1. Global Warming - Paris Accord - Methane Pledge -> obs -->
The world is experiencing an accelerated global warming
due to the accumulation of greenhouse gases (GHG)
since the industrial revolution [@us2018].
Greenhouse gas observations are critical to monitor the state of the
atmosphere, quantify present and historical emissions, and understand 
global climate change. 
During the 21th Conference of Parties (COP21), it was established the Paris 
Accord, a multilateral effort reduce greenhouse emissions
in order to limit the temperature increment of 1.5 degrees
[@rhodes20162015]. 
Methane is a greenhouse gas responsible for half of the 
temperature increase since preindustrial levels.
Furthermore, methane has a 9 years lifetime and a
global warming potential of 30 over 100 years [@epagwp], with
a current global radiative forcing of 0.650 $Wm^{-2}$ [@aggi].
Hence, in the 26 version of COP conference 
[@hunter2021glasgow], it was 
signed the Global Methane Pledge aiming reduce at least 
methane emissions 30\% from 2020 levels by 2030, with
U.S. as one of the parties [@wh]. Therefore, monitoring $CH_4$ observations,
emissions and sinks has become critical.

<!-- 2. Importance of greenhouse gases observations -->
<!-- 3. NOAA - mandate by congress - Obspack -->

The National Oceanic and Atmospheric Administration (NOAA) and its
Global Monitoring Laboratory (GML) has the mission of acquire,
evaluate and make available long-term records of atmospheric gases[^2].
To achieve that goal, GML gather own and other laboratories
data, releasing observation in a compendium named ObsPack
[@masarie2014obspack]. Specifically, the $CH_4$ ObsPack GLOBALVIEW+
is a comprehensive product consisting in observations from aircrafts, 
ships, surface stations, towers and aircores.
However, each 
ObsPack product generally contains hundreds of files, each of which 
has different sampling frequencies, hours, and hundreds of lines of 
headers. It takes time and effort to develop tools to read and process 
each ObsPack product and select observations of interest for specific 
modeling and data analysis purposes.

NOAA ObsPack data has been used to support many studies. For instance, 
the global methane budget for the year 2017 was 596 $Tgy^{-1}$,
in agreement with other studies
[@saunois2020global;@saunois2016global]
@lu2021global, characterized global methane emissions in 
between 2014 and 2017, including a comparison with Greenhouse gases Observing 
SATellite (GOSAT) data. 
@saunois2016global 
At regional scale, @lu2022methane performed another studied focused
on north america using as priors local emissions inventories. 
Furthermore, @hu2023trend presented trends and cycles of methane over the 
US.

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrfats is different than surface stations, but all the files
include concentrations and other critical fields. 
An important information is the scale of measurement, which isXX.
Cite Importance of scale and what is WMOX2014.
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome.
The `robspack` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses.

In this manuscript we present `robspack`, an R package to read,
process and plot NOAA ObsPack data. For this release, we are 
focued on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output and plots.

<!-- 4. Objective -->

# Installation

To install `robspack`, the user must have installed the R package 
remotes and run the following script. This process will install 
all the required dependencies, such as data.table, `cptcity`, an R 
package with more than 7000 color palettes, and `lubridate`, a package 
to manage time and dates [@lu;@cpt]. Then, we call the libraries
to load the function into the environment.

```{r install, eval = F}
remotes::install_github("ibarraespinosa/robspack")
library(robspack)
```

# Overview

`robspack` is a collection of function organized together to read and 
process ObsPack files [@masarie2014obspack]. The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output. In other to facilitate this task,
we are letting to the public the directory shown below. This directory 
includes a README file with all the required information and scripts 
generate process each category. The structure of the directory is:

https://github.com/ibarraespinosa/robspack/tree/main/rscripts

```{bash, eval = F}
|--"README.md"
|-- r
    |-- index.R
    |-- aircore_year.R 
    |-- aircraft_year.R
    |-- flask_non_noaa_year.R
    |-- surface_insitu_year.R
    |-- tower_insitu_year.R
    |-- inputs_inv.R
```

```{r flow, echo = FALSE, message=FALSE, fig.cap="Methane observations (ppb) by towers-insitu", out.height="50%", out.width="100%"}
knitr::include_graphics("flow/Slide1.PNG")
```

The file `index.R` creates a summary of ObsPack, generates the directories
master, receptor and obs, and store the summaries in obs directory for each
category. Then, the scripts for each category are run to generate the master
and receptor files. The next step consists in running HYSPLIT and obtaining
footprints (not in the scope of this manuscript). At last, the script
`inputs_inv.R` checks each footprint file for each category and generates the
final receptor list files. The process is described in detail in the following
example for tower insitu.

The first step consists in constructing a summary for the ObsPack. 
This is required to read the data, but also, identify agl, which is
present in some of the file names. This function returns a data.frame.
Optionally, the user can indicate a path to store the data.frame. 
obs_summary also prints a summary of the data. The second argument 
is the categories, and by default includes the categories shown below, 
to account for all the files. Then the summary data.frame contains the 
columns id as the full path to each file, name which is the name or 
relative path of the file, n which is just an id, sector such as tower,
and the column agl which indicates the agl indicated in the name of the
file if available. To read the documentation of this function, the 
user must run `?obs_summary`.

```{r libs, include = F, message=F, warning=F}
library(robspack)
library(data.table)
```

```{r}

categories <- c("aircraft-pfp","aircraft-insitu","surface-insitu",
  "aircore","surface-pfp","tower-insitu","shipboard-insitu","flask")
obs <- "../../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
index <- obs_summary(obs = obs, verbose = F)
```

```{r kable, echo = FALSE}
ta <- obs_summary(obs = obs, verbose = F, aslist = TRUE)$summary
knitr::kable(ta, format = "latex", caption = "Summary of ObsPack")
```

There are 362 files in the ObsPack directory. The printed information also 
shows the total at the bottom, as the sum of the individual file by sector. 
This is to ensure that the sum of files is equal to the total number of files 
found, shown at the top. furthermore, the printed information also shows that 
there are 136 files with the agl explicitly mentioned in the name of the file.
Sometimes we need more information about the site. For instance, what do the 
observations start and end. Then, we added the function `obs_table`, which 
calculates statistics summary of “time” and other numeric variables by file 
name, sector, site, altitude and mode. For instance, the observations in the 
site "SCT" in South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and 
"2020-12-31 23:30:00 UTC". In figure \@ref(fig:tiobs) we see the average of 
methane concentrations over each tower-insitu site. Higher concentration
are found over Russia. Over the United States (U.S.), most of
towers are found in the south to east coast.

```{r sf, echo = FALSE, message=FALSE}
library(sf)
library(ggplot2)
dft <- obs_table(index =  index, categories = "tower-insitu")
dft <- dft[stat == "mean"]
sdft <- st_as_sf(dft, coords = c("longitude", "latitude"), crs = 4326)
sdft$value <- sdft$value*1e09

```

```{r tiobs, echo = FALSE, message=FALSE, fig.cap="Methane observations (ppb) by towers-insitu"}
plot(sdft["value"], reset = FALSE, axes = TRUE, graticule = TRUE,
     pch = 16, cex = 1.5, pal = cptcity::cpt(rev = T, colorRampPalette = T, 
                                             alpha = 0.5), main = NULL)
maps::map(add = T)
```

## Functions

The `robspack` functions are shown in table 1.

Table 1. Functions and classes in `robspack`.

| Function              | Description                                       |
|-----------------------|---------------------------------------------------|
| `invfile`             | Class with `print`, `summary` and `plot` methods  |
| `obs_addltime()`      | Add local time based on metadata and longitude    |
| `obs_addtime()`       | Add UTC time                                      |
| `obs_agg()`           | Aggregates ObsPack by time                        |
| `obs_find_receptors()`| Find expected receptors and NetCDF files          |
| `obs_format()`        | Format for some columns of data.table             |
| `obs_freq()`          | Return numeric vector in intervals                |
| `obs_invfiles()`      | Construct `invfile` objects                       |
| `obs_list.dt()`       | Rbind list of data.frames with different names    |
| `obs_meta()`          | Reads ObsPack metadata                            |
| `obs_out()`           | Outersect, opposed as intersect                   |
| `obs_rbind()`         | Rbind data.frames with different names            |
| `obs_read()`          | Read files, and add metadata as columns           |
| `obs_read_csvy()`     | Read csvy file and prints yaml header             |
| `obs_roundtime()`     | Round seconds from "POSIXct" "POSIXt" classes     |
| `obs_summary()`       | Construct summary of ObsPack as a data.frame      |
| `obs_table()`         | Return a data.frame with summary of data          |
| `obs_trunc()`         | Trunc numbers with a desired number of decimals   |
| `obs_write()`         | Write CSVY to disk, YAML followed by tabulated    |

# Application for towers in situObsPack summary

## Read data

Once the summary is built, the function obs_read will read the files available in the index file previously generated. Here we selected the category "tower-insitu". The argument verbose prints which files are being read each time, by default. At the end, this function prints the total number of observations by type of altitude (agl or asl).

```{r read}
df <- obs_read(index = index,
               categories = "tower-insitu",
               verbose = FALSE)

```

We added a function to plot the data read from ObsPack. The y-axis is
the field value and the x-axis is by default time. The data illustrated
sorted by color is the field site_code, with the default number of 3 
sites. The argument pal is to define the color palette, used by the 
internally imported function cptcity::cpt.

```{r obsplot, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
obs_plot(dt = df, time = "time", yfactor = 1e9, cex = 0.5)
```

Before sub setting the data,  tower-insitu has about 
`r round(nrow(df)/1000000, 2)` million 
observations. These observations are made between 2004 and 2020. 
The identification of the altitude and type is critical, then we 
developed an approach based on the availability of data:

1. Identify agl from the name of the tile.
2. If agl not present, search fill_values used in elevation and 
transform them into NA (not available)
3. If agl is not present, agl = altitude - elevation.
4. If there are some NA in elevation, will result some NA in agl
5. A new column is added named altitude_final to store agl or asl
6. Another column named type_altitude is added to identify agl or asl.
7. If there is any case NA in altitude_final, type_altitude is 
“not available”

## Filtering

ObsPack includes global observations and sometimes we need to 
extract data for a specific region and periods of time. In this 
part we include spatial and temporal parameters to filter data. 
The year of interest is 2020, but we also included December of 
2019 and January of 2021. The spatial filtering is done by using 
the coordinates, in this case covering North America. After filtering
by space and time, we have `r round(nrow(df), 2)` million observations.

```{r filtering}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
evening <- 14:15

yy <- 2020
df <- rbind(df[year == yy - 1 & month == 12],
            df[year == yy],
            df[year == yy + 1 & month == 1])

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
```

## Time

The function obs_addtime adds time columns timeUTC, timeUTC_start 
which shows the start time of each observation and timeUTC_end which 
shows the end time for each observation. Then we need to identify the 
local time with the function add_ltime. This is important because to 
identify observations in the evening in local time for modeling 
purposes. add_ltime uses two methods, first identifying the time 
difference with utc by identifying the metadata column "site_utc2lst". 
If this information is not available, with the aircrafts for instance, 
the local time is calculated with an approximation based on longitude:

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where lt is the local time, UTC the time, and longitude is the 
coordinate. Then, the time is cut every two hours. We also identify 
the local time to select evening hours.

```{r time}
df2 <- obs_addtime(df)
df2$timeUTC <- cut(x = df2$timeUTC+3600,
                   breaks = "2 hour") |>
  as.character() |>
  as.POSIXct(tz = "UTC")
df3 <- obs_addltime(df2)
df3 <- df3[lh %in% evening]
```

Now there are `r nrow(df3)` observations. At this point we can calculate the 
averages of several columns by the cut time. The function obs_agg 
does this aggregation as shown in the following lines of code. The 
argument gby establish the function used to aggregate cols, in this 
case the function "mean" by time and altitude. Finally, we add 
local time again.

```{r agg}
df4 <- obs_agg(dt = df3,
               gby = "mean",
               cols = c("value", "latitude", "longitude", "type_altitude",
                        "dif_time", "year_end", "site_utc2lst"),
               verbose = FALSE,
               byalt = TRUE)
df5 <- obs_addltime(df4)
```

Now there are `r nrow(df5)` observations. Towers can have observations at 
different heights. Here we need to select one site with the observations 
registered at the highest height. The column with the height is named 
altitude_final and the max altitude was named max_altitude. Then, 
we print the altitudes of each site.

```{r maxalt}
df5[,
    max_altitude := max(altitude_final),
    by = site_code]
df5[,
    c("site_code",
      "altitude_final",
      "max_altitude")] |> unique()
```

## Saving master as text and csvy

Now that we have all the required information, we can save the files. 
Here, we name the data.frame as master, because it contains all the 
information. This is important because some fields can be used in the future, 
and for traceability. For convenience, time variables are transformed into 
character before writing into the disk. The separation is space ” “.

```{r fwrite, eval = T}
master <- df5
master$timeUTC <- as.character(master$timeUTC)
master$timeUTC_end <- as.character(master$timeUTC_end)
master$local_time <- as.character(master$local_time)

fwrite(master, 
       file = "tower_insitu_2020.txt",
       sep = " ")
```

The format Comma Separated Value with YAML (CSVY)[^1] consists in a 
typical CSV with a YAML header. The function obs_write includes the 
argument notes which allows adding custom notes at the header of the 
file. Below the notes, obs_write adds the output of the R function 
`str`, which provides a vertical summary of the data, known as structure.

```{r obs_write_csvy, eval = T}
obs_write_csvy(dt = master,
              notes = "tower 2020",
              out = "tower_insitu_2020.csvy")
```

To check the YAML header we read the first 38 lines of the files that 
were generated. Here we can see the column names, type of data and 
first observations. The YAML header is delimited by the characters 
"- - -" (not shown here).

```{r csvy, eval = F}
readLines("tower_insitu_2020.csvy")[1:38]
```

## Saving receptors

We need to filter some columns from the master files in a new object 
called receptors. This is needed because internally we run HYSPLIT 
[@hy] using the information from the receptors. In the case of a 
tower, we need to select observations with the highest altitude. The 
specific columns are selected as shown on the following code. We are 
selecting the ending times, because later HYSPLIT is run backwards 
based on the time of measurement, between ending and starting times. 
The columns about time are formatted to have two characters. For 
instance, the month 1, is formatted as “01”. We also need to filter 
for type_altitude equal 0, representing agl observations , or equal 
to 1, asl.

```{r receptors, eval = F}

receptor <- master[altitude_final == max_altitude,
                   c("site_code",
                     "year", "month", "day",
                     "hour", "minute", "second",
                     "latitude", "longitude",
                     "altitude_final", "type_altitude",
                     "year_end", "month_end", "day_end", "hour_end",
                     "minute_end", "second_end")]
receptor$altitude_final <- round(receptor$altitude_final)
receptor <- obs_format(receptor)

if(nrow(receptor_agl) > 0) {
  fwrite(x = receptor_agl,
         file = "paper/receptor_tower_insitu_2020_AGL.txt"),
  sep = " ")}

if(nrow(receptor_asl) > 0) {
  fwrite(x = receptor_asl,
         file = "paper/receptor_tower_insitu_2020_ASL.txt"),
  sep = " ")}

```

## Recommendation for other applications

The approach to generate receptors depends on each type of observation
and other considerations. For instance, aircraft with continuous 
observations at each second can be filtered and averaged every 20 
seconds. In that way, the footprints are still representative and it 
would not be necessary to run HYSPLIT every second. Of course, it 
depends on the application and objective of the study. For this 
manuscript, we are presenting the generation of receptors based 
on tower observations.

# Conclusion
In this manuscript we presented an robskpack, an R package to read 
and process CH4 ObsPack GLOBALVIEW+ published by the Global 
Monitoring Laboratory (GML) from the National Oceanographic and 
Atmospheric Administration (NOAA). robspack reads the text data which 
have different headers and organizes them in a common format. Then, 
this software applies calculations to filter observations by time and 
space. Finally, this software generates receptors in a suitable format 
that allows it to run HYSPLIT and generate footprints. This software 
does not provide methods to run HYSPLIT, but the user can follow the 
site https://www.ready.noaa.gov/HYSPLIT.php.

[^1]: https://csvy.org/
[^2]: https://gml.noaa.gov/about/aboutgml.html

# Acknowledgements

Funding: This project is funded by the NOAA Climate Program Office 
AC4 and COM programs (NA21OAR4310233 / NA21OAR4310234). This research was 
supported by the NOAA cooperative agreement NA22OAR4320151.

# References {-}