### 2024-04-24 examples/README_nc.Rmd
documentation and details ins NEWS.md

---
title: "R Tools for Obspack, Footprints and Receptors (rtorf)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![GitHub commit
activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)

[![R build
status](https://github.com/ibarraespinosa/rtorf/workflows/draft-pdf/badge.svg)](https://github.com/ibarraespinosa/rtorf/actions)

[![R-CMD-check](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/torf/actions/workflows/R-CMD-check.yaml)

[NOAA Obspack](https://gml.noaa.gov/ccgg/obspack/) is a collection of green house gases observations

`rtorf` only depends on `data.table`, which is basically parallel C, 
so it can be installed in any machine.

## installation

```{r, eval = F}
remotes::install_github("noaa-gml/rtorf")
```

```{r, include = T, message=F, warning=F}
library(rtorf)
library(data.table)
```

## ObsPack summary

The first step consists in constructing a summary for the ObsPack.
This is required to read the data, but also, identify `agl`, 
which is present in some of the file names.
This function returns a `data.frame`.
Optionally, the user can indicate a path to store the `data.frame`.
`obs_summary` also prints a summary of the data. The second
argument is the categories, and by default includes the categories
shown below, to account for all the files.
Then the summary `data.frame` contains 
the columns `id` as the full path to each
file, `name` which is the name or relative path of the file, `n`
just an id, `sector` such as tower, and the column `agl`
which indicates the `agl` indicated in the name of the file
if available. To read the documentation of this function,
the user must run `?obs_summary`.

>  This example reads text files but it will be updated to read NetCDF

```{r, eval = T}

cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, categories = cate)

```

There are `r nrow(index)` files in the ObsPack directory.
The printed information also shows the total at the bottom,
as the sum of the individual file by sector. This
is to ensure that the sum of files is equal to the
total number of files found, shown at the top.
furthermore, the printed information
also shows that there are 109 files with the
`agl` explicitly mentioned in the name of the file.

### Read data

Once the summary is built, the function `obs_read`
will read the files available in the
index file previously generated.
Here we selected the category "tower-insitu".
The argument verbose prints which files are being read each time,
by default.
At the end, this function prints the total number
of observations by type of altitude (agl or asl).

```{r}
df <- obs_read_nc(index = index,
               categories = "tower-insitu")

```

Sometimes we need more information about the site. For instance,
what do the observations start and end. Then, we added 
the function `obs_table`, which calculates statistics
summary of "time" and other numeric variables by file name,
sector, site, altitude and mode. For instance,
the observations in the site "SCT" in
South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
"2020-12-31 23:30:00 UTC".

```{r}
dft <- obs_table(df = df)
dft[site_name %in% unique(site_name)[sample(1:length(unique(dft$site_code)), 2)]]
```

We added a function to plot the data read from ObsPack.
The y-axis is the field `value` and the 
x-axis is by default `time`. The data illustrated sorted by
color is the field `site_code`, with the default
number of 3 sites. The argument `pal` is to define the
color palette, used by the internally imported function
`cptcity::cpt`.

```{r, tidy = TRUE, dpi=300, fig.cap="First two sites in ObsPack"}
obs_plot(dt = df, 
         time = "time", 
         yfactor = 1e9, 
         cex = 0.5)
```

Here we can see `r round(nrow(df)/1000000, 2)` million observations 
for `tower-insitu`. These observations are made between 
`r range(df$year)[1]` and `r range(df$year)[2]`. The identification
of the altitude and type is critical. The approach used here 
consists of:

1. Identify `agl` from the name of the tile.
2. If `agl` not present, search fill_values used in elevation and transform them into NA (not available)
3. If `agl` is not present, `agl = altitude - elevation`.
4. If there are some NA in elevation, will result some NA in `agl`
5. A new column is added named `altitude_final` to store `agl` or `asl`
6. Another column named `type_altitude` is added to identify `agl` or `asl`.
7. If there is any case NA in `altitude_final`, `type_altitude` is "not available"

### Filtering

ObsPack includes global observations and sometimes we need to extract data
for a specific region and periods of time.
In this part we include spatial and temporal parameters
to filter data.
The year of interest is 2020, but we also included
December of 2019 and January of 2021.
At this stage, we can apply the spatial filter
by using the coordinates.

```{r}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
evening <- 14

yy <- 2020
df <- rbind(df[year == yy - 1 & month == 12],
            df[year == yy],
            df[year == yy + 1 & month == 1])

if(any(names(df) %in% "intake_height")) {
  df <- df[intake_height < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
} else {
  df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
}

unique(df[, c("intake_height", "site_elevation", "elevation",
              "dataset_selection_tag",
              "site_name")])

```

After filtering by space and time, we have `r round(nrow(df), 2)`
million observations. Towers can have observations at different
heights. Here we need to select one site with the observations
registered at the highest height. The column
with the height is named `altitude_final` and the
max altitude was named `max_altitude`.

```{r}

if(any(names(df) %in% "intake_height")) {
dfa <- df[,
          max(intake_height),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
} else {
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
}
dfa
```

### Time

Here we need to start time columns.
The function `obs_addtime` adds time columns
`timeUTC`, `timeUTC_start` which shows the start time
of each observation and `timeUTC_end`
which shows the end time for each observation.
Then we need to identify the local time with the function
`add_ltime`. This is important because to identifying
observations in the evening in local time.
`add_ltime` uses two methods, first identify
the time difference with utc by identifying the
metadata column "site_utc2lst". If this information
is not available, with the aircrafts for instance,
the local time is calculated with an approximation
based on longitude:

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where $lt$ is the local time, $UTC$ the time,
$longitude$ the coordinate.
Then, the time is cut every two hours.
Now, we identify the local time to select evening hours.

```{r}
df2 <- obs_addtime(df)
```

"_st".

```{r}
df2 <- obs_addstime(df2)

```
solar time is not exact, then I need to cut every hour and aggregate observations
by hour. At the end I want one daily observation per day and site

# TODO ...
```{r}
df2 <- df2[hour_st %in% evening]

```

Now there are `r nrow(df3)` observations.
At this point we can calculate the averages of several columns by
the cut time. The function `obs_agg` does this aggregation as shown
in the following lines of code. The argument `gby`
establish the function used to aggregate `cols`, in this case
the function mean by time and altitude. Finally, we
add local time again.

```{r}
df4 <- obs_agg(dt = df3,
               gby = "mean",
               cols = c("value",
                        "latitude",
                        "longitude",
                        "type_altitude",
                        "dif_time",
                        "year_end",
                        "site_utc2lst"),
               verbose = FALSE,
               byalt = TRUE)
df5 <- obs_addltime(df4)
```

Now there are `r nrow(df5)` observations,
`r nrow(df3) - nrow(df5)` less observations.
Here we add the column `max_altitude` to identify the max altitude
by site.

```{r}
df5[,
    max_altitude := max(altitude_final),
    by = site_code]
df5[,
    c("site_code",
      "altitude_final",
      "max_altitude")] |> unique()
```

### Saving master as text and csvy

Now that we have all the required information, 
we can save the files. Here, we name the data.frame 
as master, because it contains all the information.
This is important because some 
fields can be used in the future, and for traceability.
For convenience,
time variables are transformed into character
before writing into the disk.
The separation is space " ".

```{r, eval = T}
master <- df5
master$timeUTC <- as.character(master$timeUTC)
master$timeUTC_end <- as.character(master$timeUTC_end)
master$local_time <- as.character(master$local_time)

fwrite(master, 
       file =  paste0(tempdir(), "/tower_insitu_2020.txt"),
       sep = " ")
```

The format Comma Separated Value with YAML (CSVY)[^3]
consists in a typical CSV with a YAML header.
The function`obs_write_csvy` includes the argument
`notes` which allows adding custom notes at the header of the file. Below the notes, `obs_write_csvy` adds 
the output of the R function `str`, which
provides a vertical summary of the data, known as
structure.

```{r, eval = T}
csvy <- paste0(tempdir(), "/tower_insitu_2020.csvy")
obs_write_csvy(dt = master,
               notes = "tower 2020",
               out = csvy
)
```

To check the YAML header we read the first 38 lines of 
the files that were generated. Here we can see
the column names, type of data and first observations.
The YAML header is delimited by the characters "---".

```{r, eval = F}
readLines(csvy)[1:38]
```

### Saving receptors

We need to filter some columns
from the master files in a new object called receptors.
This is needed because internally we run HYSPLIT [@hy]
using the information from the receptors.
In the case of a tower, we need to select 
observations with the highest 
altitude. The specific columns are selected
as shown on the following code. We are selecting the 
ending times, because later HYSPLIT is run backwards
based on the time of measurement, between ending and
starting times. The columns about time are formatted to
have two characters. For instance, the month 1,
is formatted as "01". We also need to filter for
`type_altitude` equal 0, representing `agl `observations
, or equal to 1, `asl`.

```{r, eval = F}

receptor <- master[altitude_final == max_altitude,
                   c("site_code",
                     "year",
                     "month",
                     "day",
                     "hour",
                     "minute",
                     "second",
                     "latitude",
                     "longitude",
                     "altitude_final",
                     "type_altitude",
                     "year_end",
                     "month_end",
                     "day_end",
                     "hour_end",
                     "minute_end",
                     "second_end")]
receptor$altitude_final <- round(receptor$altitude_final)
receptor <- obs_format(receptor)

if(nrow(receptor_agl) > 0) {
  fwrite(x = receptor_agl,
         file = paste0(tempdir(), "/receptor_tower_insitu_2020_AGL.txt"),
         sep = " ")}

if(nrow(receptor_asl) > 0) {
  fwrite(x = receptor_asl,
         file = paste0(tempdir(), "/receptor_tower_insitu_2020_ASL.txt"),
         sep = " ")}

```

## Application for other sectors
In this package we are sharing scripts to process other sectors
The scripts are available in the path 
`https://github.com/ibarraespinosa/rtorf/tree/main/rscripts`

## Implementation in python:

I'm currently implementing a version in python:

```{r, include=F}
data.frame(
  stringsAsFactors = FALSE,
                 R = c("fex","invfile",
                       "obs_addltime","obs_addstime","obs_addtime","obs_agg",
                       "obs_find_receptors","obs_footname","obs_format","obs_freq",
                       "obs_index","obs_invfiles","obs_list.dt","obs_meta",
                       "obs_out","obs_plot","obs_rbind","obs_read","obs_read_csvy",
                       "obs_read_nc","obs_roundtime","obs_summary",
                       "obs_table","obs_trunc","obs_write_csvy","plot.invfile",
                       "print.invfile","sr","summary.invfile"),
       description = c("File extension",
                       "Methods for objects with class 'invfile'",
                       "local hour (bsed on longitude and time)","Add solar time into obspack",
                       "Add times into obspack","Aggregates ObsPack by time",
                       "Compares expected receptors","Expected footprint name",
                       "Formatting data","return numeric vector in intervals",
                       "Summary of the ObsPack files (.txt)",
                       "Generate files to perform inverse modeling","list.dt","Read obspack metadata",
                       "outersect","Read obspack metadata","rbind obspack",
                       "Read obspack (.txt)","reads CSVY","Read obspack (.nc)",
                       "round seconds from \"POSIXct\" \"POSIXt\" classes",
                       "Summary of the ObsPack files (.txt)","Obspack Table",
                       "Trunc numbers with a desired number of decimals",
                       "Generates YAML and write data.frame",
                       "Methods for objects with class 'invfile'","Methods for objects with class 'invfile'",
                       "Extacts n last characters",
                       "Methods for objects with class 'invfile"),
            Python = c("","","","","","","","",
                       "","","OK","","","","","","","","","OK","",
                       "","","","","","","","")
) -> df

```

```{r, eval=TRUE, echo=FALSE}
knitr::kable(df)

```