### 2024-08-22 examples/aircraft-flask.Rmd
add plots to articles

---
title: "aircraft-flask"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aircraft-pfp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rtorf)
library(data.table)
```

```{r read1, eval = T}

cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "/media/sergio/ext5/obspack/obspack_ch4_1_GLOBALVIEWplus_v6.0_2023-12-01/data/nc/"
index <- obs_summary(obs = obs, 
                     categories = cate)
```

Now we read the aircraft-pfp using the function `obs_read_nc`. To this date,
`solar_time` is not included for aircraft, so we `FALSE` that argument.

```{r readnc, eval = T}
datasetid <- "aircraft-flask"
df <- obs_read_nc(index = index,
                  categories = datasetid,
                  solar_time = FALSE,
                  verbose = T)

```

Now we check the data

```{r checkdf, eval = T}
df
```

Now we can process the data. We first filter for observations
within our spatial domain:

## Checks and definitions

```{r spatial, eval = T}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
yy <- 2020

```

We check altitude, intake_height, altitude_final and elevation.
altitude_final is a column from intake_height, added to match
column from obs_read text files.

```{r checkcols, eval = T}
df[, c("altitude", "altitude_final", "intake_height", "elevation")]
```

The temporal range of data is
```{r range_year, eval = T}
range(df$year)
```

We also check for dimensions of data

```{r dim_df, eval = T}
dim(df)
```

## Filters

```{r spatial_temporal_filter, eval = T}
df <- df[year == yy]

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
dim(df)
```

## Adding time as POSIXct class

```{r add_time1, eval = T}
df <- obs_addtime(df)
df[, "timeUTC"]
```

## Cut time

now we can cut time every 20 seconds. We can chosse other frequency as well.

```{r cut_sec, eval = T}
df$sec2 <- obs_freq(x = df$second,
                     freq = seq(0, 59, 20))
df[, c("second", "sec2")]
```

## Aggregate data

now we need to add the column key_time, that it will be used
to aggregate other variables

```{r key_time, eval = T}
df$key_time <- ISOdatetime(year = df$year,
                           month = df$month,
                           day = df$day,
                           hour = df$hour,
                           min = df$minute,
                           sec = df$sec2,
                           tz = "UTC")
df[, c("timeUTC", "key_time")]

```

```{r aggregatingdata, eval = T}
df2 <- obs_agg(df, cols =  c("year",
                             "month",
                             "day",
                             "hour",
                             "minute",
                             "second",
                             "time",
                             "time_decimal",
                             "value",
                             "latitude",
                             "longitude",
                             "altitude_final",
                             "elevation",
                             "intake_height",
                             # "gps_altitude",
                             # "pressure",
                             # "pressure_altitude",
                             # "u", "v", "temperature",
                             "type_altitude"))
```
## Add local time

Now we add local time

```{r addlt, eval = T}
df3 <- obs_addltime(df2)
setorderv(df3, cols = c("site_code", "timeUTC"),
          order = c(-1, 1))
df3

```

## Master

Before generating the receptors list, we have the databe with all the required information

```{r rename, eval = T}
master <- df3

```
We may replace missing values with a nine nines. Here is commented

#master[is.na(master)] <- 999999999

We transform the time variables to character and round coordinates with 4 digits

```{r round, eval = T}
master$timeUTC <- as.character(master$timeUTC)
master$local_time <- as.character(master$local_time)
master$latitude <- round(master$latitude, 4)
master$longitude <- round(master$longitude, 4)

```

## Save master
Finally we save the master file

```{r outfile, eval = T}
out <- tempfile()

```

### txt

```{r outtxt, eval = T}
message(paste0(out,"_", datasetid, ".txt\n"))
fwrite(master,
       paste0(out,"_", datasetid, ".txt"),
       sep = " ")

```

### csv

```{r outcsv, eval = T}
message(paste0(out,"_", datasetid, ".csv\n"))
fwrite(master,
       paste0(out,"_", datasetid, ".csv"),
       sep = ",")

```

### csvy

CSVY are csv files with a YAML header to include metadata in tabulated
text files

```{r csvy, eval = T}
cat("\nAdding notes in csvy:\n")
notes <- c(paste0("sector: ", datasetid),
           paste0("timespan: ", yy),
           paste0("spatial_limits: north = ", north, ", south = ", south, ", east = ", east, ", west = ", west),
           "data: Data averaged every 20 seconds",
           paste0("altitude: < ", max_altitude),
           "hours: All",
           "local_time: if var `site_utc2lst` is not available, calculated as",
           "longitude/15*60*60 (John Miller)")

cat(notes, sep = "\n")

message(paste0(out,"_", datasetid, ".csvy\n"))
obs_write_csvy(dt = master,
               notes = notes,
               out = paste0(out,"_", datasetid, ".csvy"))
```

```{r readcsvy, eval = T}
obs_read_csvy(paste0(out,"_", datasetid, ".csvy"))
```

## Receptors

Now we can do the last step which is generating the receptor list files.
Now we filter selected columns

```{r receptors, eval = T}
receptor <- master[, c("site_code",
                       "year",
                       "month",
                       "day",
                       "hour",
                       "minute",
                       "second",
                       "latitude",
                       "longitude",
                       "altitude_final",
                       "type_altitude")]

```

We can round altitude also

```{r round_alt, eval = T}
receptor$altitude_final <- round(receptor$altitude_final)
```

Now we can format time variables with two digits

```{r formatrec, eval = T}
receptor <- obs_format(receptor,
                        spf =  c("month", "day",
                                 "hour", "minute", "second"))
```

We have a column that indicate AGL or ASL

```{r aslagl, eval = T}
receptor_agl <- receptor[type_altitude == 0]
receptor_asl <- receptor[type_altitude == 1]
```

Finally, we save the receptors

```{r save_receptors, eval = T}
if(nrow(receptor_agl) > 0) {
  message(paste0(out, "_", datasetid, "_receptor_AGL.txt"), "\n")

  fwrite(x = receptor_agl,
         file = paste0(out, "_", datasetid, "_receptor_AGL.txt"),
         sep = " ")
}

if(nrow(receptor_asl) > 0) {
  message(paste0(out, "_", datasetid, "_receptor_ASL.txt"), "\n")

  fwrite(x = receptor_asl,
         file = paste0(out, "_", datasetid, "receptor_ASL.txt"),
         sep = " ")

}

```

## Plot

Finally, we just plot some data, run it locally

```{r obs_plotsave, fig.width=7, fig.height=5, eval = T, echo = F, message=F, warning=F}
png("../man/figures/obsplot_aircraftflask.png", width = 1500, height = 1000, res = 200)
obs_plot(df3, time = "timeUTC", yfactor = 1e9)
dev.off()
```

```{r obs_plot, fig.width=7, fig.height=5, eval = F}
obs_plot(df3, time = "timeUTC", yfactor = 1e9)
```

![](https://github.com/noaa-gml/rtorf/blob/main/man/figures/obsplot_aircraftflask.png?raw=true)