### 2025-03-12 README.Rmd
fix #18 #16 #12, rename torf,

---
title: "R Tools for Observations, Receptors and Footprints (rtorf)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<img src="man/figures/logo.png" align="right" alt="" width="220" />

![GitHub commit activity](https://img.shields.io/github/commit-activity/y/noaa-gml/rtorf)
[![R-CMD-check](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/noaa-gml/rtorf/actions/workflows/R-CMD-check.yaml)
![GitHub Repo stars](https://img.shields.io/github/stars/noaa-gml/rtorf)

[NOAA Obspack](https://gml.noaa.gov/ccgg/obspack/) is a collection of greenhouse gases observations

`rtorf` only depends on `data.table` and `ncdf4`, which is basically parallel C, 
so it can be installed in any machine.

## Installation

Using git bash/powershell
```{bash installb, eval = F}
git clone https://github.com/noaa-gml/rtorf
R CMD INSTALL rtorf
```

Using remotes R package
```{r install, eval = F}
remotes::install_github("noaa-gml/rtorf")
```

```{r call, include = T, message=F, warning=F}
library(rtorf)
library(data.table)
```

## ObsPack summary

The first step consists in constructing a summary for the ObsPack.
This is required to read the data, but also, identify `agl`, 
which is present in some of the file names.
This function returns a `data.frame`.
Optionally, the user can indicate a path to store the `data.frame`.
`obs_summary` also prints a summary of the data. The second
argument is the categories, and by default includes the categories
shown below, to account for all the files.
Then the summary `data.frame` contains 
the columns `id` as the full path to each
file, `name` which is the name or relative path of the file, `n`
just an id, `sector` such as tower, and the column `agl`
which indicates the `agl` indicated in the name of the file
if available. To read the documentation of this function,
the user must run `?obs_summary`.

>  We first define the categories

```{r read1, eval = T}

cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, categories = cate)

```

For each one of these dataset ids, check the articles in the documentation

https://noaa-gml.github.io/rtorf/

## Implementation in python:

I'm currently implementing a version in python [pytorf](https://github.com/noaa-gml/pytorf):

## Special thanks to all the contributors

[![Contributors](https://contrib.rocks/image?repo=noaa-gml/rtorf)](https://github.com/noaa-gml/rtorf/graphs/contributors)