### 2025-05-28 paper/paper.Rmd
fix obs_hysplit_control when use time params instead of df. I needed

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: "R Tools for Observations, Receptors and Footprints (rtorf) for processing atmospheric observations in NOAA-GML"
tags:
  - R
  - GHG
  - NOAA
  - ObsPack
authors:
  - name: Sergio Ibarra-Espinosa
    orcid: 0000-0002-3162-1905
    affiliation: "1, 2" # (Multiple affiliations must be quoted)
  - name: Lei Hu
    orcid: 0000-0002-5624-3832
    affiliation: 2
affiliations:
 - name: Cooperative Institute for Research in Environmental Sciences, CU Boulder
   index: 1
 - name: NOAA Global Monitoring Laboratory
   index: 2
citation_author: Ibarra-Espinosa and Hu
date: 21 May 2025
year: 2025
bibliography: mybibfile.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

In this study, we present a new open-source R package `rtorf`, 
to read, process, select, and plot NOAA Observation Package (ObsPack)
data products. We use a methane ObsPack data product as an example
in this code base, but it can be easily modified to analyze ObsPack
products for other greenhouse gasses. The R package starts with 
creating a catalog of  all ObsPack files in each product. It then 
reads all files and creates one database. While reading each ObsPack 
file, it extracts site elevation and time zone information from the 
file header and calculates sampling altitude in meters above ground 
level and local time for individual events. Finally, it 
processes and selects observations for inverse modeling purposes. 
This package imports functions from data.table R package, which 
contains C bindings with parallel implementation via Open-MP 
[@dt]. rtorf provides functions 
to perform these tasks in a transparent and efficient way, 
supporting open-source communities in environmental sciences.

The world is experiencing an accelerated global warming
due to the accumulation of greenhouse gases (GHG)
since the industrial revolution [@us2018].
Greenhouse gas observations are critical to monitor the state of the
atmosphere, quantify present and historical emissions, and understand 
global climate change. 
During the 21th Conference of Parties (COP21), it was established the Paris 
Accord, a multilateral effort reduce greenhouse emissions
in order to limit the temperature increment of 1.5 degrees
[@rhodes20162015]. 
Methane is a greenhouse gas responsible for half of the 
temperature increase since preindustrial levels.
Furthermore, methane has a 9 years lifetime and a
global warming potential of 30 over 100 years [@epagwp], with
a current global radiative forcing of 0.650 $Wm^{-2}$ [@aggi].
Hence, in the 26 version of COP conference 
[@hunter2021glasgow], it was 
signed the Global Methane Pledge aiming reduce at least 
methane emissions 30\% from 2020 levels by 2030, with
U.S. as one of the initial parties [@wh]. Therefore, monitoring $CH_4$ observations,
emissions and sinks has become critical. NOAA 
ObsPack data has been used to 
support many studies. For instance, 
the global methane budget for the year 2017 was 596 $Tgy^{-1}$,
in agreement with other studies
[@saunois2020global;@saunois2016global]
@lu2021global, characterized global methane emissions in 
between 2014 and 2017, including a comparison with Greenhouse gases 
Observing SATellite (GOSAT) data. 
@saunois2016global.
At regional scale, @lu2022methane and @lei performed another studied focused
on north america using as priors local emissions inventories.

<!-- 2. Importance of greenhouse gases observations -->
<!-- 3. NOAA - mandate by congress - Obspack -->

The National Oceanic and Atmospheric Administration (NOAA) and its
Global Monitoring Laboratory (GML) has the mission of acquire,
evaluate and make available long-term records of atmospheric gases[^1].
To achieve that goal, GML gather own and other laboratories
data, releasing observation in a compendium named ObsPack
[@masarie2014obspack]. Specifically, the $CH_4$ ObsPack GLOBALVIEW+
is a comprehensive product consisting in observations from aircrafts, 
ships, surface stations, towers and aircores.
ObsPack include a descriptor named `datasetid` covering:
aircraft-pfp, aircraft-insitu, aircraft-flask, surface-insitu,
surface-flask, surface-pfp, tower-insitu, aircore, shipboard-insitu,
and shipboard-flask. ObsPack product generally contains hundreds of files, each of which 
has different sampling frequencies, hours, and attributes.
It takes time and effort to develop tools to read and process 
each ObsPack product and select observations of interest for specific 
modeling and data analysis purposes.

The NOAA ObsPack data is delivered to the public as NetCDF and 
text files [@masarie2014obspack]. The structure of the files including descriptor fields
depend on the type of file. For instance, the metadata
from aircrafts is different than surface stations, but all the files
include concentrations and other critical fields. 
Given the complexity of ObsPack format, reading and analyzing the data
can be cumbersome. The `rtorf` package provides the 
GHG science and research community a transparent and efficient tool to 
process ObsPack products for GHG modeling and analyses. 
In this manuscript we present `rtorf`, an R package to read,
process and plot NOAA ObsPack data, a software useful and needed
for the community [@R]. For this release, we are 
focused on the $CH_4$ ObsPack GLOBALVIEW+ product.
The general process consists in creating a 
summary of the ObsPack files, reading them in an iteration process, 
filtering, and generating another output and plots.

# Installation

To install `rtorf`, the user must have installed the R package 
remotes and run the following script. This process will install 
all the required dependencies, such as data.table, `cptcity`, an R 
package with more than 7000 color palettes, and `lubridate`, a package 
to manage time and dates [@lu;@cpt]. Then, we call the libraries
to load the function into the environment. `rtorf` is hosted at GitHub,
which allows the implementation of checking the package installation in a 
variety OS. To have a general view of rtorf, the reader can view the 
online diagram of the package[^2].

```{r install, eval = F}
remotes::install_github("noaa-gml/rtorf")
```

# Overview

`rtorf` is a collection of function organized together to read and 
process ObsPack files. The general process 
consists in create a 
summary of the ObsPack files, reading them in an iteration process, 
filter and generating another output. As $CH_4$ ObsPACKGLOBALViewplus 5.1,
the product used in this manuscript, includes dataid, we produced
a guide for each of of them available at https://noaa-gml.github.io/rtorf.
Then, in this manuscript we present the processing of `aircraft-insitu`.
The obspack product in this case is `obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08`.

## Summary

We first call the libraries `rtorf` and `data.table`.  Most of objects returned
by `rtorf` are of class `data.table`. Then, we define the datasetid to be 
identified in the name of the files inside the directory with data. 
This process print a summary of the data and if the logical argument verbose
is `TRUE`, print the file being read in each iteration, default is `FALSE`.

```{r}
library(rtorf)
library(data.table)
cate = c("aircraft-pfp","aircraft-insitu", "aircraft-flask", 
         "surface-insitu", "surface-flask", "surface-pfp", "tower-insitu", 
         "aircore", "shipboard-insitu","shipboard-flask")

obs <- "Z:/obspack/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, 
                     categories = cate)
```

Once the index of file is built, we can read each file. As we are directing to the
nc directory in ObsPack, with NetCDF files inside, we use the function 
`obs_read_nc`. This function dumps the NetCDF information into a data.table 
with `long` format [@Silge2016]. As the global attributes is attributes in
the NetCDF would result in a data.table with too many columns, we used
the argument `att` equals to FALSE, which is default.
In ground-based datasetid the `solar_time` array is available.
This is useful to select specific observations, for more information check
the site documentation. At the moment, this array is not available for aircraft
observations, hence we select `FALSE`. In this case, we select `verbose` equal
to `TRUE`, to see the name of the files being read.

```{r read_air}
datasetid <- "aircraft-insitu"
df <- obs_read_nc(index = index, categories = datasetid,
                  att = FALSE, solar_time = FALSE, verbose = TRUE)
```

The resulting `data.table` contains `r ncol(df)` columns and 
`r nrow(df)` observations. 
Furthermore, the size of `data.table` is `r format(object.size(df), unit = "Gb")`.
The data includes observations between `r range(df$year)[1]` and 
`r range(df$year)[2]`. Now we can define some parameters to filter our data,
like the year 2020 and spatially data below 8000 meters above sea level (masl) and
focused over north America. Nevertheless, it can be modified to any region.

```{r}
df <- df[year == 2020 & altitude_final < 8000 & latitude < 80 &
           latitude > 10 & longitude < -50 & longitude > -170]
```

Now we have a `data.table` that contains `r ncol(df)` columns and 
`r nrow(df)` observations. The size of `data.table` is 
`r format(object.size(df), unit = "Gb")`. Sometimes the data can be alaready
filtered every 20 seconds or for a different period of time.
However, raw data can be availalb ein a second-by-second basis. 
Under this case we may need to aggregate datah. In this example, we can add a column
of time in format "POSIXct" and cut the seconds every 20 seconds.
Usually, aircraft observations every 1 second. Then,
We can simplify the data by calculating averages every 20 seconds.
We perform this task by cutting time very 20 seconds. Then,
we add a new column with the mandatory name `key_time`, which 
will be used to aggregate data with "POSIXct" class, but every 20 seconds.

```{r}
df <- obs_addtime(df)
df$sec2 <- obs_freq(x = df$second,
                    freq = seq(0, 59, 20))
df$key_time <- ISOdatetime(year = df$year, month = df$month, day = df$day,
                           hour = df$hour, min = df$minute, sec = df$sec2,
                           tz = "UTC")
df[1, c("timeUTC", "key_time")]
```

now we can aggregate the data using the function `obs_agg`. The argument cols
indicate which columns will be averaged. Then, we add local time with the 
function `obs_addltime` and we re order the data.table.

```{r}
df2 <- obs_agg(df, 
               cols =  c("year", "month", "day", "hour", "minute",
                         "second", "time", "time_decimal",  "value",
                         "latitude", "longitude", "altitude_final",
                         "pressure", "u", "v", "temperature",
                         "type_altitude"))
df3 <- obs_addltime(df2)
setorderv(df3, cols = c("site_code", "timeUTC"),
          order = c(-1, 1))
```

## Solar or local time

Identifying the local time is important for atmospheric reasons. Sometimes
we need observations when the Planetary Boundary Layer is high,
so that the concentrations are well distributed, in genera 
around 2:00pm, when planetary boundary layer is higher. 
In `rtorf` we use an hierarchical approach
based on the availability of critical information.
Basically, if the solar time array is available, we use
the function `obs_addstime`. In the negative case
`rtorf` searches for the metadata `site_utc2lst` to convert
UTC time to local. Finally, in the absence of the mentioned data,
we calculate an approximation of the local time
using the geographical coordinates, as :

$$
lt = UTC + longitude/15 * 60 * 60
$$
Where $lt$ is the local time, $UTC$ the time,
$longitude$ the coordinate.

## Plots

Now we have the data processed and ready to be exported. `rtorf` includes 
a number of functions to save the data as tabulated format in text, csv and 
CSVY[^3] are csv files with a YAML header. This functions can be seen in the documentation.
In this last part of the manuscript we will show some visualizations. We included
a function named `obs_plot` which plots data in long format using R base functions.
Here we see data for the month of March 2020. This useful function allows to plot several
sites and prints the x-axis range.

```{r, eval = F}
obs_plot(df3[month == 3], time = "timeUTC", yfactor = 1e9, 
         type = "b", xlab = "UTC time", ylab = expression(CH[4]~ppb))
```

Finally, we show some vertical profiles for the months of January and March of 2020.
We can see how during March of 2020 methane concentrations are lower than
January. This may be due the implementation of Lockdowns [@espinosa2023covid].
A manuscript focused on the impact of COVID-19 on methane emissions
will be submitted soon.

```{r, eval = F}
x <- df3
x$ch4 <- x$value*1e+9
obs_plot(x, time = "ch4",  y = "altitude_final",  colu = "month", 
         type = "b",  xlab = expression(CH[4]~ppb),  ylab = "altitude (m)")
```

```{r, echo=FALSE, out.width="90%", fig.cap="a) Time series, b) Monthly observations by altitude", fig.show='hold',fig.align='center'}
knitr::include_graphics(c("paper_files/obsplot_ts.png",
                          "paper_files/obsplot_alt.png"))
```

# HYSPLIT

`rtorf` also provides functions to run HYSPLIT (Hybrid Single-Particle Lagrangian Integrated Trajectory) model [@stein2015noaa], through `obs_hysplit_control`, `obs_hysplit_ascdata`
and `obs_hysplit_setup`. These function were designed to be used inside programs
and run using `rslurm` for parallel processing[^4]. This capability is particularly valuable for in-depth analysis of atmospheric observations, helping to interpret measurement data in the context of air mass histories and contributing to more robust emission quantifications and atmospheric model evaluations.

## Future work

We are currently porting `rtorf` to python into a package named `pytorf`[^5].

[^1]: https://gml.noaa.gov/
[^2]: https://gitdiagram.com/noaa-gml/rtorfml
[^3]: https://csvy.org/
[^4]: https://noaa-gml.github.io/rtorf/articles/hysplit.ht
[^5]: https://github.com/noaa-gml/pytorf

# Acknowledgements

This project is funded by the NOAA Climate Program Office AC4 and COM programs (NA21OAR4310233 / NA21OAR4310234). This research was supported by NOAA cooperative agreement NA22OAR4320151. Also, thanks to Arlyn Andrews, John Miller, Kenneth Schuldt, Kirk Thoning and Andy Jacobson from NOAA GML.

```{r, echo=FALSE, message=FALSE }
invisible(system("cp paper.md ../"))
```

# References